### 2024-11-19 LakeMetabolism_project/project_22/reports/project_report1_2022.Rmd
First push

---
title: "Report data-processing sondes-project 2022"
author: "Emanuel Mauch"
date: "2023-07-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This is a report of the data-processing steps of the Greenland sonde data of 2022. I adapted Moritz' code
from his paper (https://doi.org/10.1002/ecy.3371). I have cleaned all the raw datasets, stored them as .txt files (folder "Raw_sonde_2022_3"), merged them with RStudio, processed the merged dataset and plotted some figures of the absolute values. The code was adapted from Moritz' file "make_data.R", using his "method_packages.R". All the files I've used to make this report are in the "project_22" folder that I have uploaded.

# Steps done so far
## Step 0: Clean raw datafiles in Excel, store them as .txt files
I cleaned all the raw .csv files with Excel (e.g. reassembling split files, fixing date bugs which lead to wrong .txt file conversions etc.). They can be found in the folder "Raw_sonde_2022_3".

## Step 1: set up R-script
```{r}
rm(list= ls())

setwd("~/ZIVI_EAWAG/project_22")

source("~/ZIVI_EAWAG/project_22/Moritz_Luehrig_paper_stuff/methods_packages.R")
library(GGally)

sonde_key = fread( "~/ZIVI_EAWAG/project_22/ponds_sonde_key.txt", header=T)
print(head(sonde_key))
```
## Step 2: merge all raw data files into a single big one
```{r, warning = F}
path = "~/ZIVI_EAWAG/project_22/Raw_sonde_2022_3"
filenames <- list.files(path=path, pattern=".txt")
files<-data.table(NULL)

system.time(for(i in filenames){
  filepath <- file.path(path,paste(i,sep=""))
  dummy1<-fread(filepath,
                colClasses = "character" ,
                sep = "\t",
                header = FALSE,
                fill=TRUE,
                encoding ="UTF-8")
  rownumb<-dummy1[V1 %like% "(MM/DD/YYYY)", which=TRUE]+1
  colname<-dummy1[V1 %like% "(MM/DD/YYYY)"]
  colname<-gsub(" ", "",(gsub("[[:punct:]]", "",c(lapply(colname,as.character)))))
  colname<-unlist(strsplit(colname,"\t"))
  dummy2<-dummy1[rownumb:.N]
  setnames(dummy2,colname)
  dummy2[,(colname[c(3,5:length(dummy2))]):=lapply(.SD, as.numeric),.SDcols=colname[c(3,5:length(dummy2))]]
  dummy2$Sonde<-substr(i, 1,7)
  dummy2$Source_file<-i
  files<-rbindlist(list(files, dummy2), fill=TRUE)
})

all <- files

# save intermediate stage 1 (FULL RAW DATA, NO REMOVAL)
fwrite(all, "~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_1.txt", sep="\t")
print(head(all,3))
```

## Step 3: add exp. design, phases and format time
```{r}
all <- fread("~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_1.txt", sep="\t")

# format date & time
all$TimeHHMMSS<-substr(all$TimeHHMMSS,1,5) #gsub("\\D+", "", cleaning$TimeHHMMSS)
all$DateMMDDYYYY<-mdy(all$DateMMDDYYYY)
all$Date_time<-ymd_hm(paste(all$DateMMDDYYYY, all$TimeHHMMSS))
setnames(all,1:2,c("Date","Time"))

# add fractions of day since start
# all = all[order(Date_time, Sonde)]
all[,Time_seq:=round(((as.numeric(Date_time))-
                        (as.numeric(Date_time)[1]))/86400,7)]

# merge with key
all<-merge(all,sonde_key, by = "Sonde")

# reorder and revalue
setorderv(all, c("Pond", "Sonde", "Date_time"), c(1, 1, 1))

# select columns to keep
all<-all[,c("Pond","Sonde", "Treatment", "Time_seq", "Date_time",
            "ChlorophyllRFU", "ConduScm" , "BGAPCRFU", "ODOmgL", "pH", "fDOMRFU", "SpConduScm", "TempC")]

# rename
setnames(all, c("Pond", "Sonde", "Treatment", "Time_seq", "Date_time",
                "Chlorophyll_RFU", "Cond_uScm", "BGAPC_RFU", "ODO_mgL", "pH", "fDOM_RFU", "SpCond_uScm", "Temp_C"))

# debug: rm duplicate value
all = unique(all, by=c("Pond", "Time_seq"))

fwrite(all, "~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_2.txt", sep="\t")
print(head(all,3))
```

# Step 4.1: remove outliers
The outlier plots are in the folder "outliers".

```{r, eval=FALSE}
all <- fread("~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_2.txt", sep="\t")

save_path = "outliers/"

for(i in  c("Chlorophyll_RFU", "BGAPC_RFU", "ODO_mgL")){
  all[,  (i) := lapply(.SD[,..i],
                       outlier_removal,
                       window=48,
                       threshold=10,
                       plot=T,
                       return="removed",
                       save=T,
                       path=save_path,
                       name=paste0(i,"_",Pond)),
      by = c("Pond")]
}

for(i in c("fDOM_RFU")){
  all[,  (i) := lapply(.SD[,..i],
                       outlier_removal,
                       window=96,
                       threshold=10,
                       plot=F,
                       return="removed",
                       save=T,
                       path=save_path,
                       name=paste0(i,"_",Pond)),
      by = c("Pond")]
}

for(i in c("Temp_C", "SpCond_uScm", "Cond_uScm")){
  all[,  (i) := lapply(.SD[,..i],
                       outlier_removal,
                       window=96,
                       threshold=20,
                       plot=F,
                       return="removed",
                       save=T,
                       path=save_path,
                       name=paste0(i,"_",Pond)),
      by = c("Pond")]
}

# cut-off "loose ends" on both sides -> all same start and same end of measurements
all <- subset(all, Date_time>"2022-06-23 22:00:00" & Date_time<"2022-06-29 05:00:00")

# add fractions of day since start
# all = all[order(Date_time, Sonde)]
all[,Time_seq:=round(((as.numeric(Date_time))-
                        (as.numeric(Date_time)[1]))/86400,7)]

# ommit date-time (just needed for cutting the dates)
all<-all[,!c("Date_time")]

# save
fwrite(all, "~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_3.1.txt", sep="\t")
```

## Step 4.2: remove anomalies
We inspected the data for anomalies and remove them here - for details see 
Russo, S., M. Lührig, W. Hao, B. Matthews, and K. Villez. 2020. 
Active learning for anomaly detection in environmental data. 
Environmental Modelling & Software 134:104869.

I didn't do this step, since I would need some time to refresh my python skills, 
but if you think it is worth, I will manage.

## Step 5: add light and precipitation data
I do not have this data available.

## Step 6: intercept correction
This uses the results form a cross correlation survey to adjust the 
sondes for "off factory differences, i.e. remove any intercept variation
among the data.

I do not have the sonds correlation factors.

## Step 7: save
```{r, eval=FALSE}
## order and remove rownames
all = all[order(Pond, Time_seq), ]

rownames(all)<-NULL

fwrite(all, "~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")
```

```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")
print(head(all,3))
```

## Step 8: check completeness
```{r}
all <- fread("~/ZIVI_EAWAG/project_22/intermediate_processing_steps/ponds_sonde_data_intermediate_3.1.txt", sep="\t")
all$Date_time <- as.POSIXct((all$Time_seq*86400), origin="2022-06-23 20:00:00", tz="UTC")

all$Date <-date(all$Date_time)

# N per date
all[, N := uniqueN(Date), by = Sonde]
complete <- setDT(data.frame(unclass(table(all$Date, all$Sonde))), keep.rownames=T)
colnames(complete)[1]<-"Date"
fwrite(complete, "~/ZIVI_EAWAG/project_22/data/sonde_data_completeness.txt", sep="\t")
print(complete)
```

## Step 9: calculate daily averages
```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")
summary(all)
all[,Time_seq:=as.integer(Time_seq)]

parameters = c("Chlorophyll_RFU", "Cond_uScm", "BGAPC_RFU", "ODO_mgL", "pH",
               "fDOM_RFU", "SpCond_uScm", "Temp_C")
groups = c("Time_seq", "Pond", "Sonde", "Treatment")

all.summ = all[,lapply(.SD, m_r, 2), by=groups, .SDcols = parameters]

fwrite(all.summ, "~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_daily_avg.txt", sep=",")

table(all.summ$Pond, all.summ$Treatment)
print(head(all.summ,3))
```

## Step 10: hourly averages
```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")

parameters = c("Chlorophyll_RFU", "Cond_uScm", "BGAPC_RFU", "ODO_mgL", "pH",
               "fDOM_RFU", "SpCond_uScm", "Temp_C")

all$Date_time <- as.POSIXct((all$Time_seq*86400), origin="2022-06-23 20:00:00", tz="UTC")

all$Time_seq = as.integer(all$Time_seq)
all$Hour <- as.numeric(substr(as.character(all$Date_time),12,13))

groups = c("Time_seq", "Hour","Pond", "Treatment")

all.summ = all[,lapply(.SD, m_r, 2), by=groups, .SDcols = parameters]

fwrite(all.summ, "~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_hourly_avg.txt", sep=",")
print(head(all.summ,3))
```

## Step 11: 10-minute averages
I thought that maybe the hour interval is too large for our time series,
so maybe a 10-minute interval could be more appropriate (or just plot¨
the absolute values).

```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")

parameters = c("Chlorophyll_RFU", "Cond_uScm", "BGAPC_RFU", "ODO_mgL", "pH",
               "fDOM_RFU", "SpCond_uScm", "Temp_C")

all$Date_time <- as.POSIXct((all$Time_seq*86400), origin="2022-06-23 20:00:00", tz="UTC")

all$Time_seq = as.integer(all$Time_seq)
all$Hour <- as.numeric(substr(as.character(all$Date_time),12,13))
all$Ten_min <- as.numeric(substr(as.character(all$Date_time),15,15))

groups = c("Time_seq", "Hour", "Ten_min", "Pond", "Treatment")

all.summ = all[,lapply(.SD, m_r, 2), by=groups, .SDcols = parameters]

fwrite(all.summ, "~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_ten_min_avg.txt", sep=",")
print(head(all.summ,3))
```

## Step 12: calculate overall averages of individual ponds
```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")

parameters = c("Chlorophyll_RFU", "Cond_uScm", "BGAPC_RFU", "ODO_mgL", "pH",
               "fDOM_RFU", "SpCond_uScm", "Temp_C")
groups = c("Pond","Treatment", "Sonde")

all.summ = all[,lapply(.SD, m_r, 2), by=groups, .SDcols = parameters]

fwrite(all.summ, "~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_overall_avg.txt", sep=",")

table(all.summ$Pond, all.summ$Treatment)
```

## Step 13: control plots
```{r}
all<-fread("~/ZIVI_EAWAG/project_22/data/ponds_sonde_data_all.txt")

all$Date_time <- as.POSIXct((all$Time_seq*86400), origin="2022-06-23 22:00:00", tz="UTC")

table(all$Pond, all$Treatment)
```

### Chlorophyll RFU
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
# Chloropyll_RFU
p1 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p2 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p1, p2, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# Chloropyll_RFU
```

### BGAPC RFU
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
# BGAPC RFU
p5 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p6 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p5, p6, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```

### ODO mg/L
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
# ODO mg/L
p3 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p4 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p3, p4, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```

### Cond uS/cm
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
# Cond uS/cm
p7 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p8 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p7, p8, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```

### Speficic Cond uS/cm
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
p9 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p10 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p9, p10, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```

### fDOM RFU
```{r, echo=FALSE, fig.height=8, fig.width = 9, warning = FALSE}
# fDOM RFU
p11 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p12 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p11, p12, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```

# 14: Appendix
## Version and packages used to generate this report:
```{r, echo=FALSE}
cat(paste(Sys.time(), Sys.timezone(), "\n"))
sessionInfo()
```
## Code used to generate control plots:
```{r, eval=FALSE}
# Chloropyll_RFU
p1 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p2 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=Chlorophyll_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p1, p2, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# ODO mg/L
p3 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p4 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=ODO_mgL, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p3, p4, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# BGAPC RFU
p5 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p6 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=BGAPC_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p5, p6, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# Cond uS/cm
p7 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p8 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=Cond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p7, p8, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# Speficic Cond uS/cm
p9 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p10 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=SpCond_uScm, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p9, p10, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))

# fDOM RFU
p11 = ggplot(subset(all, Pond=="B1P1")) + theme_light() +
  geom_line(aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P2"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P3"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B2P3"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P1"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P2"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B1P1', 'B2P2', 'B3P3', "B2P3", "B3P1", "B3P2"),
                     values=c('B1P1'="gray17", 'B2P2'='dodgerblue3', 'B3P3'='magenta4',
                              "B2P3"="olivedrab4", "B3P1"="chocolate", "B3P2"="midnightblue")) +
  ggtitle("Fish")

p12 = ggplot(subset(all, Pond=="B2P4")) + theme_light() +
  geom_line(aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P0"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="B3P4"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL152"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  geom_line(data = subset(all, Pond=="ERL85"), aes(y=fDOM_RFU, x=Date_time, color = Pond)) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%m-%d") +
  scale_color_manual(breaks=c('B2P4', 'B3P0', 'B3P4', "ERL152", "ERL85"),
                     values=c('B2P4'="gray17", 'B3P0'='dodgerblue3', 'B3P4'='magenta4',
                              "ERL152"="olivedrab4", "ERL85"="chocolate")) +
  ggtitle("No Fish")

plot_grid(p11, p12, nrow = 2, rel_heights =  c(0.229,0.2,0.2,0.32))
```