### 2024-12-05 LakeMetabolism_project/Synthesis_21_22_23/report/manuscript.Rnw
Work on manuscript

%%
%% Camera-ready submissions do not need line numbers, and
%% should have this option removed.
%%

\documentclass[fleqn,11pt,lineno]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Template for manuscript}

\author[1]{Author One}
\author[1]{Author Two}
\author[2,3]{Author Three}
\author[1]{Author Four}
\affil[1]{Author one affiliation}
\affil[2]{Author two affiliation}
\affil[3]{Author three affiliation}

\corrauthor[1]{Author Four}{email@address}

\keywords{Keyword1; Keyword2; Keyword3}

\begin{abstract}
This is the abstract.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% R specifications
<<include = FALSE>>=
# knitr
library(knitr)
opts_chunk$set(fig.path = 'plots/p', 
               echo = FALSE, 
               results = 'hide',
               fig.height = 4.5,
               fig.align = "center",
               cache = TRUE)

# packages
require(xtable) # generating latex tables from datasets
require(data.table) # data wrangling
require(readxl) # importing excel tables

@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Materials and Methods}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Author contributions}
...

\section*{Institutional review}
...

\section*{Data availability} 
...

\section*{Funding}
...

\section*{Acknowledgments}
...

\section*{Conflicts of interest}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\beginsupplement
\renewcommand\figurename{Supplementary Figure}
\renewcommand\tablename{Supplementary Table}

\section*{Appendix}

\subsection{Study design and setup} \label{subsec:design}

In 2019, 12 glacial lakes ($N =$ 12) were chosen for an experiment near Narsarsuaq,
Greenland (61.1567\textdegree{}N, -45.4254\textdegree{}E).
These glacial lakes vary in area (0.21 to 1.82 hectares) and maximum
depth (2 to 8 m), but are all clustered within a few
kilometers of each other. All lakes were fishless at the beginning of the experiment.
Six lakes were subsequently introduced with three-spined sticklebacks
(\textit{Gasterosteus aculeatus}) from nearby lakes.
Lake B1P1, B2P2, and B3P3 were introduced with \textit{Gasterosteus aculeatus} from a single
population (lake L26, 61.253333\textdegree{}N, -45.529141\textdegree{}E),
while lake B2P3, B3P1, and B3P2 were introduced with \textit{Gasterosteus aculeatus} from
two populations (lake L26, 61.253333\textdegree{}N, -45.529141\textdegree{}E
and lake ERL33, 61.118369\textdegree{}N, -45.580845\textdegree{}E).
The remaining six lakes B1P4, B2P4, B3P0, ERL85, ERL122, and ERL152 were
used as fishless control. For the purpose of this study, the origin
of the introduced \textit{Gasterosteus aculeatus} is of minor importance,
as they all originate from the same area (Supplementary Table~\ref{tab:lakes}).

In 2021, 2022, and 2023, all 12 lakes were monitored over several days.
For that purpose, EXO2 multiparameter sondes were installed
(YSI, Yellow Springs, OH, USA), tracking ecosystem parameters with high
frequency (2-minute intervals in 2021 and 2022, 5-minute intervals in
2023 with the exception of ERL122, which was monitored in 15-minute intervals).
For the purpose of this study, only dissolved oxygen and temperature measurements
yielded from these sondes are relevant. The sensors were situated at a water depth
of approximately 1-1.5 m in each lake. All optical sensors were
wiped clean before every measurement with a built-in wiper.
The monitoring period was 16 September-24 September in 2021, 22 June-3 July in 2022,
and 22 June-17 July in 2023.

<<laketable, results = "asis", echo=FALSE>>=
# Import lakemaster file
lakes <- read_excel("../../Organisation/GL_Lake_Master.xlsx")

# Extract relevant ponds
lakes <- lakes[c(18:23, 73:75, 78, 28, 83),]

# Extract columns
lakes <- lakes[,c("Lake", "Lat", "Long", "Areaha", "Altitude", "Depth")]

# Rename lakes
lakes$Lake <- c("B1P1", "B2P2", "B3P3", "B2P3", "B3P1", "B3P2", "B1P4", "B2P4",
                "B3P0", "ERL85", "ERL122", "ERL152")

# Rename
setnames(lakes, c("Lake", "Latitude (°N)", "Longitude (°E)", "Area (hectare)",
                  "Altitude (m)", "Maximum Depth (m)"))

# Create treatment variable
lakes$Treatment <- c("Fish","Fish","Fish","Fish","Fish","Fish",
                   "No Fish","No Fish","No Fish","No Fish","No Fish","No Fish")

# Reorder
lakes <- lakes[,c("Lake", "Treatment", "Latitude (°N)",
                  "Longitude (°E)", "Altitude (m)", "Area (hectare)",
                  "Maximum Depth (m)")]

# Make a latex table from dataset
lakes <- xtable(lakes, align = c("l","l","l","l","l","l","l","l"),
                digits = c(0,0,0,5,5,0,2,2),
                caption = "Lakes included in the experiment, along with treatment
                and general characteristics.", label = "tab:lakes")

# Print
print(lakes, include.rownames = FALSE, caption.placement = "top", booktabs = TRUE)
@

\subsection{Data sources} \label{subsec:source}

Dissolved oxygen and water temperature measurements were yielded from EXO2 multiparameter
sondes, as described in section~\ref{subsec:design}. For the purpose
of estimating ecosystem metabolism, wind and irradiation data were yielded
from the QAS\_L automated weather station near Narsarsuaq, Greenland \citep{metadata}.

\subsubsection*{Weather data}

\subsection{Data preparation}

\subsection{Statistical analysis}

\subsection{Implementation}

\end{document}