### 2024-12-06 LakeMetabolism_project/Synthesis_21_22_23/report/manuscript.Rnw
Work on  manuscript

%%
%% Camera-ready submissions do not need line numbers, and
%% should have this option removed.
%%

\documentclass[fleqn,11pt,lineno]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue,
    }

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Template for manuscript}

\author[1]{Author One}
\author[1]{Author Two}
\author[2,3]{Author Three}
\author[1]{Author Four}
\affil[1]{Author one affiliation}
\affil[2]{Author two affiliation}
\affil[3]{Author three affiliation}

\corrauthor[1]{Author Four}{email@address}

\keywords{Keyword1; Keyword2; Keyword3}

\begin{abstract}
This is the abstract.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% R specifications
<<include = FALSE>>=
# knitr
library(knitr)
opts_chunk$set(fig.path = 'plots/p', 
               echo = FALSE, 
               results = 'hide',
               fig.height = 4.5,
               fig.align = "center",
               cache = TRUE)

# packages
require(xtable) # generating latex tables from datasets
require(data.table) # data wrangling
require(readxl) # importing excel tables

@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Materials and Methods}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Author contributions}
...

\section*{Institutional review}
...

\section*{Data availability} 
...

\section*{Funding}
...

\section*{Acknowledgments}
...

\section*{Conflicts of interest}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\beginsupplement
\renewcommand\figurename{Supplementary Figure}
\renewcommand\tablename{Supplementary Table}

\section*{Appendix}

\subsection{Study design and setup} \label{subsec:design}

In 2019, 12 glacial lakes ($N =$ 12) were chosen for an experiment near Narsarsuaq,
Greenland (61.1567\textdegree{}N, -45.4254\textdegree{}E).
These glacial lakes vary in area (0.21 to 1.82 hectares) and maximum
depth (2 to 8 m), but are all clustered within a few
kilometers of each other. All lakes were fishless at the beginning of the experiment.
Six lakes were subsequently introduced with three-spined sticklebacks
(\textit{Gasterosteus aculeatus}) from nearby lakes.
Lake B1P1, B2P2, and B3P3 were introduced with \textit{Gasterosteus aculeatus} from a single
population (lake L26, 61.253333\textdegree{}N, -45.529141\textdegree{}E),
while lake B2P3, B3P1, and B3P2 were introduced with \textit{Gasterosteus aculeatus} from
two populations (lake L26, 61.253333\textdegree{}N, -45.529141\textdegree{}E
and lake ERL33, 61.118369\textdegree{}N, -45.580845\textdegree{}E).
The remaining six lakes B1P4, B2P4, B3P0, ERL85, ERL122, and ERL152 were
used as fishless control. For the purpose of this study, the origin
of the introduced \textit{Gasterosteus aculeatus} is of minor importance,
as they all originate from the same area (Supplementary Table~\ref{tab:lakes}).

In 2021, 2022, and 2023, all 12 lakes were monitored over several days.
For that purpose, EXO2 multiparameter sondes were installed
(YSI, Yellow Springs, OH, USA), tracking ecosystem parameters with high
frequency (2-minute intervals in 2021 and 2022, 5-minute intervals in
2023 with the exception of ERL122, which was monitored in 15-minute intervals).
For the purpose of this study, only dissolved oxygen and temperature measurements
yielded from these sondes are relevant. The sensors were situated at a water depth
of approximately 1-1.5 m in each lake. All optical sensors were
wiped clean before every measurement with a built-in wiper.
The monitoring period was 16 September-24 September in 2021, 22 June-3 July in 2022,
and 22 June-17 July in 2023.

<<laketable, results = "asis", echo=FALSE>>=
# Import lakemaster file
lakes <- read_excel("../../Organisation/GL_Lake_Master.xlsx")

# Extract relevant ponds
lakes <- lakes[c(18:23, 73:75, 78, 28, 83),]

# Extract columns
lakes <- lakes[,c("Lake", "Lat", "Long", "Areaha", "Altitude", "Depth")]

# Rename lakes
lakes$Lake <- c("B1P1", "B2P2", "B3P3", "B2P3", "B3P1", "B3P2", "B1P4", "B2P4",
                "B3P0", "ERL85", "ERL122", "ERL152")

# Rename
setnames(lakes, c("Lake", "Latitude (°N)", "Longitude (°E)", "Area (hectare)",
                  "Altitude (m)", "Maximum Depth (m)"))

# Create treatment variable
lakes$Treatment <- c("Fish","Fish","Fish","Fish","Fish","Fish",
                   "No Fish","No Fish","No Fish","No Fish","No Fish","No Fish")

# Reorder
lakes <- lakes[,c("Lake", "Treatment", "Latitude (°N)",
                  "Longitude (°E)", "Altitude (m)", "Area (hectare)",
                  "Maximum Depth (m)")]

# Make a latex table from dataset
lakes <- xtable(lakes, align = c("l","l","l","l","l","l","l","l"),
                digits = c(0,0,0,5,5,0,2,2),
                caption = "Lakes included in the experiment, along with treatment
                and general characteristics.", label = "tab:lakes")

# Print
print(lakes, include.rownames = FALSE, caption.placement = "top", booktabs = TRUE)
@

\subsection{Data sources} \label{subsec:source}

Dissolved oxygen and water temperature measurements were yielded from EXO2 multiparameter
sondes, as described in section~\nameref{subsec:design}. For the purpose
of estimating ecosystem metabolism, wind and irradiation data were yielded
from the Programme for Monitoring of the Greenland Ice Sheet (PROMICE), providing automatic
weather station data \citep{Fausto:2021}.
The dataset named \texttt{QAS\_L\_hour.csv} originating from the QAS\_L automated weather station
near Narsarsuaq was downloaded from a
\href{https://dataverse.geus.dk/citation?persistentId=doi:10.22008/FK2/IW73UU}{public database}
provided by PROMICE \citep{metadata}.

\subsection{Data preparation}

The data preparation for estimating lake metabolism was done separately for each monitoring
period (16 September-24 September 2021, 22 June-3 July 2022, and 22 June-17 July 2023).

\subsubsection*{Sonde data}

First, the raw sonde data from each lake were imported and merged into
one dataset for a given monitoring period. The variables needed for lake metabolism
estimation were water temperature (\textdegree{}C) and
dissolved oxygen (mg/L). Next, an outlier analysis was performed by removing
values higher than three times the median absolute deviation of all values in a
sliding window of one day window size \citep{Luerig:2021}.
After outlier removal, the time series of water temperature and dissolved oxygen
were subsequently investigated for larger data gaps due to potential lack of sensor
measurements over certain periods.

To ensure overlapping estimation of lake metabolism, the data was subsequently
cut to achieve equal starting and ending time points for all lakes during a given
observation period. Due to large data gaps during the period of 2022, we chose
to keep as much complete days as possible in order to estimate the maximum
amount of days, rather than selecting the longest common monitoring period
for all lakes.

After cutting, missing values were replaced by a
weighted moving average, where the weights of observations around the
central value decrease exponentially \citep{Moritz:2017}.
For larger data gaps spanning over several hours, we chose
to not estimate lake metabolism for the affected days, rather than imputing these values.

\subsubsection*{Weather data}

\subsection{Statistical analysis}

\subsection{Implementation}

\end{document}